\chapter{Preface}

This document describes the RISC-V Xvpfloat extension unprivileged architecture.

The document contains the version 1.1.0 of the RISC-V Xvpfloat ISA extension.


%The changes in this version of the document include:
%\vspace{-0.2in}
%\begin{itemize}
%\parskip 0pt
%\itemsep 1pt
%\end{itemize}

\section*{Preface to Document Version 20210618-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Add first major version of the Xvpfloat ISA extension.
\end{itemize}

\section*{Preface to Document Version 20210625-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Change the order of summary bits.
\item Add Permitted Hardware Configurations.
\item Describe all instructions.
\end{itemize}

\section*{Preface to Document Version 20210702-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Change the order of environment fields.
\item Change src field of PCVT.P.(H/F/D) to rs2.
\item Rename PADD0 to PRND.
\item Add details on padding when loading/storing at high precision.
\end{itemize}

\section*{Preface to Document Version v0.9-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Hide unused header bits.
\item Add memory consitency/coherency details.
\item Replace MBB by BIS.
\end{itemize}

\section*{Preface to Document Version v0.9.1-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Fix an error in MV\_SUM description of PMV.X.P (only 4 summary bits).
\item Define handling of subnormals, NaNs and Infs.
\item Define exceptions.
\item Add Rhea implementation details (WIP).
\end{itemize}

\section*{Preface to Document Version v0.9.2-draft}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Clarification of overflow of internal registers.
\item Add first details on system registers, to be completed in a further version.
\item Hide debug and perf monitoring sections.
\end{itemize}

\section*{Preface to Document Version v1.1.3}

The changes in this version of the document include:
\vspace{-0.2in}
\begin{itemize}
\parskip 0pt
\itemsep 1pt
\item Change in the pcvt.p.x instruction format.
\item Fix description of store.vpx instructions.
\item Fix exception part: we don't check BIS, we check the mantissa size.
\end{itemize}
